Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae

### 2021-01-31 referee_answers.txt

2. Most analysis of GW signals is based on two-dimensional models, and
the test set used in the present study is also comprised of 2D
simulations. Recent comparisons of signals from 2D and 3D simulations
show that the signal amplitude from 2D simulations is considerably
higher than in 3D simulations (e.g., Andresen et al. (2017), MNRAS,
468, 2032; see also Mezzacappa et al. 2020). I would like the authors
to comment on this, and its potential impact on the results presented
in the present paper. Would this just lower the upper limit of the
detectability distance? Are there other crucial differences between
signals from 2D and 3D simulations (e.g., the low frequency signal
from the spiral SASI mode) that could compromise the accuracy of the
model derived here? Please discuss this.

A. The referee is correct. A reduction in the amplitude of the GW signal
would reduce the detectability distance. Regarding the model we put forward
in this paper, it is indeed possible that the physical effects that only 
appear in 3D CCSN simulations could alter the relationship between the 
mode frequencies and the PNS properties that we have used in this paper. 
This work, thus, is far from providing the final word on the detectability
prospects of GW signals from PNS evolution, but it certainly serves the 
purpose of offering a promising proof-of-concept and a framework upon which 
to build a better model. We plan to use data from 3D simulations to improve 
our model and to study the differences in the relationships among 1D, 2D and 
3D simulations. We have added a comment on this important caveat in the conclusion.

"Moreover, we have used 2D CCSN simulations
that are known to produce higher amplitude GW signals
than 3D CCSN simulations [60, 61]. This fact and the
low frequency standing accretion shock instability that
appears only in 3D simulations may reduce further the
performance of the method."

@@

About the model for log-sigma, we tested only 2 models: a linear and a 
quadratic. We found that the best model is the latter according to the AIC 
criterion.  Higher order models cause that the iterative fitting procedure 
does not converge in a satisfactory manner, which is warned by the authors 
of the lmvar package.

Figure 1 shows that overall the model (for r and log sigma) fits 
quite well the data set.

We have rewritten the corresponding paragraph to clarify this.

"For the mean of the ratios $r_i$ , the cubic model has the
best fitting amongst polynomials of lower degrees, and
for the mean of the log standard deviation $\rm log \sigma_i$ , the
quadratic model has the best fit compared to a polynomial
of degree one. The models were compared according
to the Akaike information criterion. The estimated coefficients
are given in Table II and the data and fit of the
final model including 95\% confidence bands are displayed
in Figure 1."

### 2021-01-31 referee_answers.txt

A. R_{PNS} is defined as the radius where the density becomes lower than
10^12 gr/cm^3. This definition has been added in the text.

5. Can you elaborate on the motivation for the degrees of the
polynomials in Eqs. (1) and (2)? Why is r_i of degree 3 and log
sigma_i of degree 2?

A: Figure 1 gives an idea of the relationship between frequencies and 
ratios, which is apparently quadratic.  However, we tested a bunch of 
models, namely, linear, quadratic, cubic, and some other combinations.  
We assessed their performance through the AIC criterion, getting the 
cubic as the best model. This is explained in the article.

Regarding the model for log-sigma, we tested only 2 models: a linear and a 
quadratic. We found that the best model is the latter according to the AIC 
criterion.  For higher order models, the iterative fitting procedure 
does not converge which is also mentioned in a warning by the authors 
of the lmvar package.

Figure 1 shows that overall the model (for r and log sigma) fits the data set quite well.

### 2021-02-01 referee_answers.txt

We thank the refereee for his/her supportive comments and provide
answers below.

----------------------------------------------------------------------
Report of the Referee -- DY12631/Bizouard
----------------------------------------------------------------------

1. As far as I can tell, the model assumes that the GW signal is
dominated by g-modes originating from the PNS, which is consistent
with results reported from recent CCSN simulations from some groups.
However, other groups, using three-dimensional simulations, attribute
the dominant component to f- and g-modes (Radice et al. (2019), ApJL,
876, L9) or p- and g-modes (Mezzacappa et al. (2020), Phys. Rev. D,
102, 023027). Can the authors comment on this. In particular, discuss
the possible impact on model predictions if the signal is hybrid in
nature.

A. We believe that a discussion of the nature of the modes is out of 
the scope of this paper. The nature of the mode visible in the time-frequency
evolution do not affect our analysis. The only assumption is that there is 
an excited mode (whatever its nature) whose frequency scales with the properties 
of the proto-neutron star following a universal relation. It is true that
the ratio used (M/R^2) is inspired by the presence of a g-mode and we and this
interpretation is discussed in our previous work. In Torres-Forne 
et al. Phys. Rev. Lett. 123, 051102 (2019) fits to the various modes (f, p, 
and g) that appear in the spectrograms of a number of CCSN simulations were 
discussed, and universal relations for the frequency evolution were proposed.
The best fit for the dominant feature found in all simulations was that 
corresponding to what was identified as a g-mode. We agree with the referee 
that uncertainties remain on the actual nature of the various modes identified 
in simulations performed by various groups. Further work is certainly necessary 
to clarify this issue, not only regarding three-dimensional simulations but also 
regarding the development of improved perturbative models for mode identification 
that incorporate 3D effects and rotation in a more accurate way that existing 
approaches. The current uncertainty in the identification of the modes at work 
during the evolution of a PNS, while must definitely be removed, is of no 
importance for the message that this paper conveys.

We have modified slightly the text in page 2 to clarify this:

"(different authors may have slightly different naming convention)"
->
"This same mode has been interpreted in some works as a f or a p- mode [e.g. 29, 48, 49]. 
However, this does not affects our analysis, which only uses the relation between the 
frequency of the main mode emitting in GWs (regardless of its nature) and the properties 
of the PNS."

@@

2. Most analysis of GW signals is based on two-dimensional models, and
the test set used in the present study is also comprised of 2D
simulations. Recent comparisons of signals from 2D and 3D simulations
show that the signal amplitude from 2D simulations is considerably
higher than in 3D simulations (e.g., Andresen et al. (2017), MNRAS,
468, 2032; see also Mezzacappa et al. 2020). I would like the authors
to comment on this, and its potential impact on the results presented
in the present paper. Would this just lower the upper limit of the
detectability distance? Are there other crucial differences between
signals from 2D and 3D simulations (e.g., the low frequency signal
from the spiral SASI mode) that could compromise the accuracy of the
model derived here? Please discuss this.

A. The referee is correct. A reduction in the amplitude of the GW signal
would reduce the detectability distance. Regarding the model we put forward
in this paper, it is indeed possible that the physical effects that only 
appear in 3D CCSN simulations could alter the relationship between the 
mode frequencies and the PNS properties that we have used in this paper. 
Sotani & Takiwaki 2020a,b, compared simulations with different dimensionality 
(1D, 2D and 3D) and found no differences for a limited set of models (this
is mentioned in our work in the introduction). This work, thus, is far from 
providing the final word on the detectability prospects of GW signals from PNS 
evolution, but it certainly serves the purpose of offering a promising 
proof-of-concept and a framework upon which to build a better model. We plan 
to use data from 3D simulations to improve our model and to study the differences 
in the relationships among 1D, 2D and 3D simulations. We have added a comment on 
this important caveat in the conclusion.

@@

3. Are the 2D simulations new (i.e., computed for the present paper),
or have they been published elsewhere? If they have been published
elsewhere, please provide citation. Were any of the 2D simulations
constrained to spherical symmetry in the inner core to alleviate the
CFL condition? If so, could this restriction potentially suppress
modes that would otherwise have been excited by instabilities in the
PNS?

A. The simulations are new, except the s20 which was used in Torres-Forne 
et al. Phys. Rev. Lett. 123, 051102 (2019). We avoid the severe time step 
restriction near the centre by applying a coarsening scheme to the hydrodynamic 
and neutrino data.  Where the lateral grid width is smaller than the radial 
one (in the innermost ~ 10 km), we combine in theta-direction several grid cells 
of the original grid to larger cells.  The resulting effective time step criterion 
is identical to the radial one, i.e., corresponding to several 100 m.  That way, 
we save up to two orders of magnitude in computing time.  This method does not 
imply a reduction to spherical symmetry close to the origin, but maintains the 
two-dimensional nature and allows for angular flows such as the modes generating 
the GW signal.

4. How do you define R_{PNS} in the 1D and 2D simulations? Is it the
radius for which the density achieves a specific value? Please clarify
this.

A. R_{PNS} is defined as the radius where the density becomes lower than
10^11 gr/cm^3. This definition has been added in the text.

### 2021-02-01 introduction.tex
Working in the reply

{In this work we introduce a method to infer PNS properties, namely a combination of the mass and radius, using GW information. For this purpose we have developed an algorithm to  
  extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} (different authors may have slightly different naming convention). Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), \textcolor{red}{where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{12}$]{$\rm g/cm^3$}}.
%  where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.
Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

>>>>

{In this work we introduce a method to infer PNS properties, namely a combination of the mass and radius, using GW information. For this purpose we have developed an algorithm to  
  extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in~\citep{Torres:2019b} 
 %(different authors may have slightly different naming convention)
Next, we use the universal relations obtained by~\citep{Torres:2019b}, based on a set of 1D simulations, to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), \textcolor{red}{where $M_{\rm PNS}$ is the mass of the PNS and $R_{\rm PNS}$ is the radius where the density is lower than \unit[$10^{11}$]{$\rm g/cm^3$}}.
%  where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.
\textcolor{red}{This same mode has been interpreted in some works as a f or a p-mode
\citep[e.g.][]{Sotani:2016,Radice:2018usf,Mezzacappa:2020lsn}. 
However, this does not affects our analysis, which 
only uses the relation between the frequency of the main mode emitting in GWs (regardless
of its nature) and the properties of the PNS.}    
Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.
